### 2019-10-08 defesa/tex/chapter/e3p.tex
wip

One of the most useful techniques when dealing with complicated functions is approximation. They appear in various methods whenever a derivative or integral needs to be calculated or for example, in our case, when the roots of a function need to be determined. In general, one has a function $f$ that is part of a family of functions $\mathcal{A}$ and wants to select a simpler function $f^*$ from a set of functions $\mathcal{A^*}$, such that $f^*$ is close enough to $f$ \cite[p.~3]{powell}. For this problem, the approximation of $\xi(\theta)$ on the interval $[0, \pi)$ is considered. The approximation set of functions is going to be the set of $N$-degree Chebyshev polynomials which the roots can be found through determining the eigenvalues of a $N$ by $N$ matrix.

@@

We refer to $T_n : [-1, 1] \mapsto [-1, 1]$ as the $n$-degree Chebyshev polynomial of the first kind:

@@

Although, its definition can be extended, so its domain is the whole real line, It also respects the following recurrence relation:

\begin{equation}
\begin{array}{ll}
T_0(x) = 1& \\
T_1(x) = x& \\
T_n(x) = 2xT_{n-1}(x) - T_{n-2}(x)&n = 2, 3, \dots
\end{array}
\end{equation}

@@

Polynomial interpolation is a form of approximating a function by a polynomial of degree $N$ that passes through $N+1$ points. In fact, this polynomial is unique and it is determined by Lagrange's formula:

\begin{equation}
\sum_{j=0}^{N} \dfrac{\prod_{k \neq j}^{N+1} (x-x_k)}{\prod_{k \neq j}^{N+1} (x_j-x_k)}.
\end{equation}

>>>>

One of the most useful techniques when dealing with complicated functions is approximation. They appear in various methods whenever a derivative or integral needs to be calculated or for example, in our case, when the roots of a function need to be determined. In general, one has a function $f$ that is part of a family of functions $\mathcal{A}$ and wants to select a simpler function $f^*$ from a set of functions $\mathcal{A^*}$, such that $f^*$ is close enough to $f$ \cite[p.~3]{powell}. For this problem, the approximation of $\xi(\theta)$ on the interval $[0, \pi)$ is considered. The approximation set of functions is going to be the set of $n$-degree Chebyshev polynomials which the roots can be found through determining the eigenvalues of a $n$ by $n$ matrix.

@@

We refer to $T_n : [-1, 1] \mapsto [-1, 1]$ as the $n$-degree Chebyshev polynomial of the first kind, and it is defined as follows:

@@

It is important to mention that this definition can be extended to the whole real line. Using some trigonometric identities, $T_n$ can also be expressed as a recurrence relation:

\begin{equation}
T_n(x) = 2xT_{n-1}(x) - T_{n-2}(x).
\end{equation}

An important property worth bringing up is that Chebyshev polynomials are orthogonal and form a basis for the polynomial space. This implies that any $p_n$ of degree up to $n$ can be expressed as a truncated Chebyshev series:

\begin{equation}\label{eq:chebseries}
p_n(x) = \sum_{j=0}^{n} a_j T_j(x).
\end{equation}

One of the greatest qualities of Chebyshev polynomials is its numerical stability. \citeonline{gautschi:1979} showed that the matrix that maps polynomials onto its coefficients written in the power form\footnote{$p_n$ is in the power form or the monomial form if it can be written as $p_n=\sum_{j=0}^{n}a_jx^j$} has a condition number that grows exponentially with $n$. On the other hand, the matrix that converts polynomials to the Chebyshev basis as \autoref{eq:chebseries}, has a linear condition number bounded by $\sqrt{2}n$.

@@

Polynomial interpolation is a form of approximating a function by a polynomial of degree $n$ that passes through $n+1$ chosen points. In fact, this polynomial is unique and it is determined by Lagrange's formula:

\begin{equation}\label{eq:lagrange}
f_n(x) = \sum_{j=0}^{n} f(x_j)\dfrac{\prod_{k \neq j}^{n+1} (x-x_k)}{\prod_{k \neq j}^{n+1} (x_j-x_k)},
\end{equation} 
with $f$ being the function to be approximated and $f_n$ the unique $n$ degree polynomial that passes through $\{(x_j, f(x_j)): j=0, 1, \dots n\}$. Because of the uniqueness of interpolant polynomials, there is a direct link between the quality of an approximation and the points chosen to interpolate. As a matter of fact, depending on the points one chooses, even increasing the degree of the interpolation makes the approximation worsen. This is known as Runge's phenomenon and an example can be seen in \citeonline[p.~37]{powell} where uniformly spaced points are chosen to interpolate the function $f(x) = (1+x^2)^{-1}$ on the interval $[-5, 5]$. That is where Chebyshev interpolation comes in, instead of $n+1$ arbitrary points, the interpolation is done on the $n+1$ roots of $T_n$, which are also known as Chebyshev Nodes:

\begin{equation}
x_j = \cos{\left(\pi\dfrac{2n-2j+1}{2n}\right)},
\end{equation}
for $j=1, \dots, n$. If the domain of the function to be interpolated is in another range other than $[-1, 1]$, let us say $[a, b]$, then a transformation can be done to map it to the Chebyshev Nodes' domain:

\begin{equation}
\hat{x_j} = \frac{a+b}{2} + \frac{b-a}{2}x_j.
\end{equation}

Then, the Chebyshev interpolation of a function $f: [a, b] \mapsto \R$ can be determined using Lagrange's formula and the points $\hat{x}_1, \dots, \hat{x}_n$. 
As it was mentioned, finding the roots of a polynomial written in the monomial form can be done by determining the eigenvalues of a so-called companion matrix. For small $n$ this works fine, however, converting the polynomial obtained by \autoref{eq:lagrange} to the power form, as $n$ grows, becomes a very ill-conditioned problem.

An alternative method can be found in \citeonline{boyd:2013}. It turns out that the Chebyshev interpolation can be calculated directly as a truncated Chebyshev series in $\bigO(n^2)$. Also, given a polynomial written in the Chebyshev basis, a $n\times n$ matrix can be constructed, such that its eigenvalues are the roots of that polynomial. Therefore, the whole process of interpolating and finding the roots can be done using only Chebyshev polynomials which have great numerical stability.
